### 2015-11-17 section/0_intro.tex
finish revising introduction

Performance bug detection tools use static or dynamic analysis to identify
code regions that match specific inefficiency patterns 
\cite{Alabama,CARAMEL, Cachetor,Xu:2010:FLD:1806596.1806617,Dufour:2008:STC:1453101.1453111, Xu:2009:GFP:1542476.1542523, Xu:2010:DIC:1806596.1806616,IsilDillig.PLDI15}. 
Unfortunately, these tools are not designed for and are consequently
unsuitable for performance diagnosis.
They are not designed to provide \textit{coverage} for a wide
variety of real-world problems; their analysis is not guided by
performance failure symptom, and hence are at disadvantage in terms of diagnosis
\texttt{accuracy}; dynamic detection tools often lead to 
10X slowdowns or more \cite{Cachetor,Xu:2010:FLD:1806596.1806617,Alabama}, 
not ideal in terms of \textit{performance}.

Recently, progress has been made on statistical debugging for
performance diagnosis \cite{SongOOPSLA2014}.
Statistical debugging, an approach widely used for diagnosing
functional failures \cite{liblit03,liblit05,tarantula1}, can
accurately identify control-flow constructs that are most correlated with
the performance problem by comparing runs with and without problematic 
performance.
Unfortunately, for loop-related 
performance problems, which contribute to two thirds of
real-world performance problems studied in previous work 
\cite{SongOOPSLA2014,PerfBug}, statistical debugging is not very effective.
Although it can identify the root-cause loop, it does not provide any 
information regarding why the loop is inefficient and hence is not very helpful
in fixing the performance problem.

@@

Figure~\ref{fig:GCC27733} shows a real-world performance bug in GCC. 
Function \texttt{synth\_mult} is used to compute the best algorithm for 
multiplying \texttt{t}. It is so time consuming that developers tried 
speeding it up through memoization ---
hash-table
\texttt{alg\_hash} is used to remember which \texttt{t} has been processed
in the past and what is the processing result. 
Unfortunately, a small mistake in the type declaration of hash-table
entry \texttt{alg\_hash\_entry} causes memoization not to work when \texttt{t}
is larger than the maximum value type-\texttt{int} can represent.
As a result, under certain workload, \texttt{synth\_mult} conducts a lot of
redundant computation for same \texttt{t} values recursively, leading to
huge performance problems.  In practice, developers know the problem is inside
\texttt{synth\_mult} very early. However, since this function is 
complicated, it took them several weeks to figure out the root cause. 
If a tool can tell them not only which loop\footnote{Recursive functions are
handled similarly as loops in this paper.}
is the root cause but also why a loop is inefficient (i.e., lot of redundant
computation and the need for better memoization in this example), 
the diagnosis and bug fixing
process would be much easier for developers.

Clearly, more research is needed to improve the state of the art of performance
diagnosis. Diagnosis techniques that can provide better
\textit{coverage} and \textit{accuracy} for common performance problems,
especially loop-related performance problems, with good
performance is well desired.

\subsection{Contributions}
This paper presents a tool, \Tool, that can help effectively diagnose
inefficient loop problems with good coverage, accuracy, and performance.

\Tool targets the most common type of performance problems, inefficient
loops, according to previous
empirical studies \cite{SongOOPSLA2014,PerfBug}.
\Tool also targets a challenging problem. Although
statistical debugging \cite{SongOOPSLA2014}
can help identify suspicious loops that are highly correlated with 
user-perceived performance problems,
it provides no information about whether and why 
the suspicious loop is inefficient,
not to mention providing fix suggestions. Although various
bug-detection techniques \cite{Cachetor,Alabama,CARAMEL} can help detect
loop-related performance bugs, they all suffer from generality problems,
covering only a specific type of loop problems each; the dynamic
tools \cite{Cachetor,Alabama} also suffer from performance
problem, imposing 10X slowdown or more.

@@

First, figuring out a taxonomy for the root causes of common inefficient loops.
Such a taxonomy is the prerequisite to developing a general and accurate
diagnosis tool. Guided by a thorough study of 45 real-world inefficient
loop problems, we come up with a hierarchical root-cause taxonomy for
inefficient loops. 
%Specifically, we consider \textit{resultless} and
%\textit{redundancy} as two main categories of root causes for inefficient
%loops, representing loops that spend effort without producing
%any results or producing redundant results. They are further divided into
%sub-categories based on the granularity of the resultless or redundant work. 
Our empirical study shows that our taxonomy is general enough to cover
common inefficient loops, and also specific enough to help developers understand
and fix the performance problem. We will present more details 
in Section \ref{sec:study}.

Second, building a tool(kit) \Tool that can automatically and accurately
identify the fine-grained root-cause type of a suspicious loop, 
together fix-strategy suggestions. 
We achieve this following several principles:

\begin{itemize}
\item Focused checking. 
Different from performance-bug detection tools that blindly checks the whole
software, \Tool focuses on performance symptoms and only check inefficient
loop candidates identified by existing tools, a statistical debugging
tool in our current prototype \cite{SongOOPSLA2014}. 
This focused study is crucial for \Tool to achieve high
accuracy.

@@

\item Static-dynamic hybrid analysis to balance performance and accuracy.
As we will see, static analysis alone cannot accurately identify 
inefficiency root causes, especially because some inefficiency problems only
happen under specific workload. However, pure dynamic analysis will 
cause too large 
run-time overhead. Therefore, we take a hybrid approach to achieve both
performance and accuracy goals.
\end{itemize}

@@

We have evaluated \Tool on 18 real-world performance problems. 
Evaluation results show that \Tool can accurately identify detailed
root cause for all benchmarks and provide correct fix-strategy suggestion for
16 benchmarks. All of these are achieved with low run-time overhead.

>>>>

Performance bug detection tools use static or dynamic analysis to identify
code regions that match specific inefficiency patterns 
\cite{Alabama,CARAMEL, Cachetor,Xu:2010:FLD:1806596.1806617,Dufour:2008:STC:1453101.1453111, Xu:2009:GFP:1542476.1542523, Xu:2010:DIC:1806596.1806616,IsilDillig.PLDI15}. 
Unfortunately, these tools are not designed for and are consequently
unsuitable for performance diagnosis.
They are not designed to provide \textit{coverage} for a wide
variety of real-world problems; their analysis is not guided by
performance failure symptom, and hence are at disadvantage in terms of diagnosis
\textit{accuracy}; dynamic detection tools often lead to 
10X slowdowns or more \cite{Cachetor,Xu:2010:FLD:1806596.1806617,Alabama}, 
not ideal in terms of \textit{performance}.

Recently, progress has been made on statistical debugging for
performance diagnosis \cite{SongOOPSLA2014}.
This approach compares runs with and without problematic performance, and
accurately identifies control-flow constructs, such as a branch $b$ or 
a loop $l$, that are most correlated with
the execution slowness.%, referred to as slowness-predictor.
Unfortunately, this approach is \textit{not} effective
for loop-related 
performance problems, which contribute to two thirds of
real-world performance problems studied in previous work 
\cite{SongOOPSLA2014,PerfBug}. It cannot
tell whether and how loop $l$ is inefficient 
and hence is not very helpful
in fixing performance problems.

@@

Figure~\ref{fig:GCC27733} shows a real-world performance bug in GCC. 
Function \texttt{mult\_alg} is used to compute the best algorithm for 
multiplying \texttt{t}. It is so time consuming that developers used a
hash-table
\texttt{alg\_hash} to remember which \texttt{t} has been processed
in the past and what is the result. 
Unfortunately, a mistake in the type declaration of hash-table
entry \texttt{hash\_entry} makes the memoization useless for large \texttt{t},
severely hurting performance.
In reality, developers easily figured out that function
\texttt{mult\_alg} is most correlated with the execution slowness, but
spent several weeks to finally figure out \texttt{mult\_alg} is in fact
inefficient.
%If a tool can tell them not only which loop\footnote{Recursive functions are
%handled similarly as loops in this paper.}
%is the root cause but also why a loop is inefficient (i.e., lot of redundant
%computation and the need for better memoization in this example), 
%the diagnosis and bug fixing
%process would be much easier for developers.

Clearly, more research is needed to improve the state of the art of performance
diagnosis --- better diagnosis
coverage, better diagnosis accuracy, and low run-time
overhead for common performance problems,
especially loop-related performance problems.

\subsection{Contributions}
This paper presents a tool, \Tool, that can help effectively diagnose
inefficient loop problems, the most common type of performance problems
\cite{SongOOPSLA2014,PerfBug}, with good coverage, accuracy, and performance.

@@

First, figuring out a root-cause taxonomy for inefficient loops.
Such a taxonomy is the prerequisite to developing a general and accurate
diagnosis tool. Guided by a thorough study of 45 real-world inefficient
loop problems, we come up with a hierarchical root-cause taxonomy for
inefficient loops that are
%Specifically, we consider \textit{resultless} and
%\textit{redundancy} as two main categories of root causes for inefficient
%loops, representing loops that spend effort without producing
%any results or producing redundant results. They are further divided into
%sub-categories based on the granularity of the resultless or redundant work. 
general enough to cover
common inefficient loops, and also specific enough to help developers understand
and fix performance problems. More details are
in Section \ref{sec:study}.

Second, building a tool(kit) \Tool that can automatically and accurately
identify whether and how a suspicious loop is inefficient, 
together with fix-strategy suggestions. 
We achieve this following several principles:

\begin{itemize}
\item Focused checking. 
Different from performance-bug detectors that blindly checks the whole
software, \Tool focuses on loops that are
most correlated with performance symptoms. 
This focus is crucial for \Tool to achieve both high
accuracy and high coverage.

@@

\item Static-dynamic hybrid analysis.
As we will see, static analysis alone cannot accurately identify 
inefficiency root causes, especially because some inefficiency problems only
happen under specific workload. However, pure dynamic analysis will 
cause too large 
run-time overhead. Therefore, we take a hybrid approach to achieve both
performance and accuracy goals.
\end{itemize}

@@

We evaluated \Tool on 18 real-world performance problems. 
Evaluation results show that \Tool can accurately identify detailed
root cause for all benchmarks and provide correct fix-strategy suggestion for
16 benchmarks. All of these are achieved with low run-time overhead.
